Barriers to a summer fire regime in northern prairies

### 2023-06-26 SpringSummerSeverity/talk/NAPC2023.tex
finish talk

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{FiraSans}
\usepackage{graphicx} 
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{adjustbox}
\usepackage{soul}
\usepackage{caption}
	\captionsetup[figure]{	labelformat=empty, 
					font=scriptsize,
					labelfont=scriptsize, 
					justification=raggedleft}

@@

@@

\begin{frame}{The Western fire regime concept}
	\begin{center}
		\begin{figure}
			\includegraphics[width=0.9\linewidth]{figs/FireRegimeModulatorsNAPC} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{Rx fire in the Northern Great Plains}
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/region_map-1.pdf} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{Rx fire in the Northern Great Plains}
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/PBGmaps} 
		\end{figure}
	\end{center}
\end{frame}

@@

@@

@@

\begin{frame}{What a difference a few days makes! } 
	\begin{columns}
		\begin{column}{0.5\textwidth}
				\begin{center}
				\begin{figure}
					%\caption{5 May 2018}
					\includegraphics[width=0.9\linewidth]{figs/GoodBurn} 
				\end{figure}
			\end{center}
		\end{column}
		\begin{column}{0.5\textwidth}  
			\begin{center}
				\begin{figure}
					%\caption{16 May 2018}
					\includegraphics[width=1\linewidth]{figs/StreeterFire_17}  
				\end{figure}
			\end{center}
		\end{column}
	\end{columns}
\centering
\adjustbox{max height=\dimexpr\textheight-5.5cm\relax,
	max width=1\textwidth}{
	\begin{tabular}{ccc}
		\toprule
		5 May 2018 & \textbf{Date} & 16 May 2018 \\
		Weather & \textbf{Wind} & \\
		Fuels & \textbf{RH} & Yes \\ \bottomrule
	\end{tabular}
}
\end{frame}

\begin{frame}{Barriers to summer Rx fire } 
	All opportunities and limitations fit within fire regime concept
	\begin{itemize}
		\item Biophysical 
		\item[] \emph{mostly, \alert{too wet}}
		\begin{itemize}
			\item High live moisture fuel content (\emph{photosynthesis})
			\item High dead moisture fuel content (\emph{humidity})
			\item Poor convection/smoke dispersal (\emph{humidity})
		\end{itemize}
	\item[]
		\item Social 
		\item[] \emph{mostly, \alert{too dry}}
		\begin{itemize}
			\item Local burn restrictions
			\item Control issues
			\end{itemize}
	\end{itemize}

@@

@@

\begin{frame}{Fire, grazing, and the pre-European landscape}
\alert{Pyric herbivory}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
	\item Recent burns have high-quality forage
	\item[]
	\item Grazers focus on burned areas
	\item[]
	\item Vegetation grows up in other areas...
		\begin{itemize}
			\item providing habitat
			\item fueling future fires
		\end{itemize}
	\end{itemize}
\end{column}
\begin{column}{0.55\textwidth}  
\begin{center}
\begin{figure}
 \includegraphics[width=1\linewidth]{figs/ghd}

\begin{frame}{Disturbance-driven vs. inherent heterogeneity}
\begin{center}
\begin{figure}
 \includegraphics[width=1\linewidth]{figs/StreeterFire_17} 
 \end{figure}
\end{center}
\end{frame}

\begin{frame}{Testing PBG 2.0 in the Northern Great Plains } 
	\begin{columns}
		\begin{column}{0.6\textwidth}
			\begin{itemize}
				\item \alert{Streeter (Central Grasslands)}
				\begin{itemize}
					\item \textbf{Compares grazing systems}
					\item 4 pastures x 3 systems; cattle
					\item \emph{Poa pratensis}-invaded mixed-grass prairie
					\item 46 cm (18") annual rainfall
				\end{itemize}
				\item \alert{Hettinger}
				\begin{itemize}
					\item \textbf{Compares cattle \& sheep}
					\item 3 x cattle, 3 x sheep PBG
					\item Mostly non-native, cool-season grasses
					\item post-CRP pasture stands
					\item 40 cm (16") annual rainfall
				\end{itemize}
			\end{itemize}
		\end{column}
		\begin{column}{0.45\textwidth}  
			\begin{center}
				\begin{figure}
					\includegraphics[width=1\linewidth]{figs/BobPasture}  
				\end{figure}
			\end{center}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Mopping up} 
\alert{Prescribed fire has direct benefits to livestock production}
\begin{columns}
\begin{column}{0.55\textwidth}
\begin{itemize}
	\item Livestock focus grazing in recently-burned patches
	\item[]
	\item Burned areas have better forage
	\item[]
	\item Fire has positive or neutral effects on soil
	\item[]
	\item Patch-burning enhances conservation outcomes, too
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}  
\begin{center}
\begin{figure}
\includegraphics[width=1\linewidth]{figs/mop_up} 
 \end{figure}
\end{center}
\end{column}
\end{columns}
\end{frame}

>>>>

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{FiraSans}
\usepackage{textcomp} %% put this in your preamble
\usepackage{graphicx} 
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{adjustbox}
\usepackage{soul}
\usepackage{caption}
	\captionsetup[figure]{	labelformat=empty, 
					font=scriptsize,
					labelfont=scriptsize, 
					justification=raggedleft}

@@

\begin{frame}{For every season: Burn, burn, burn...}
		\vspace{-3em}	
	\begin{columns}

		\begin{column}{0.5\textwidth}
			\begin{itemize} 
				\item Conventional R\textsubscript{x} fire conducted during primarily dormant season
				\item[]
				\item Increased interest in burning during non-dormant season
				\item[]
				\begin{itemize}
					\item Awareness of pre-colonial fire regimes
					\item Diversify management
				\end{itemize}

			\end{itemize}

		\end{column}
		\begin{column}{0.5\textwidth}  
			\begin{center}
				\begin{figure}
					\includegraphics[width=1\linewidth]{figs/StreeterRxFire}

				\end{figure}
			\end{center}
		\end{column}
	\end{columns}

	%\alert{As with most components of fire regime, \textbf{seasonality is a construct}}
\end{frame}

\begin{frame}{For every season: Burn, burn, burn...}

	\begin{columns}
		\begin{column}{0.5\textwidth}
			\begin{itemize} 
				\item Conventional R\textsubscript{x} fire conducted during primarily dormant season
				\item[]
				\item Increased interest in burning during non-dormant season
				\item[]
				\begin{itemize}
					\item Awareness of pre-colonial fire regimes
					\item Diversify management
				\end{itemize}

			\end{itemize}

		\end{column}
		\begin{column}{0.5\textwidth}  
			\begin{center}
				\begin{figure}
					\includegraphics[width=1\linewidth]{figs/StreeterRxFire}

				\end{figure}
			\end{center}
		\end{column}
	\end{columns}
\alert{As with most fire regime parameters, \textbf{seasonality is a construct}}
\end{frame}

@@

\begin{frame}{Controls over fire behavior and effects}
	\begin{center}
		\begin{figure}
			\includegraphics[width=0.9\linewidth]{figs/FireRegimeModulatorsNAPC} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{Case study: R\textsubscript{x} fire in the Northern Great Plains}
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/region_map-1.pdf} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{R\textsubscript{x} fire in the Northern Great Plains}
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/PBGmaps} 
		\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}{Remotely-sensed burn severity products}
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/severity_gg-1} 
			\caption{An example from Dunn Ranch Prairie (McGranahan et al. 2022 \emph{Land})}
		\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}{Barriers to summer R\textsubscript{x} fire } 
	All opportunities and limitations fit within fire regime concept
	\begin{itemize}
		\item Biophysical 
		\item[] \emph{mostly, \alert{too wet}}
		\begin{itemize}
			\item High live moisture fuel content (\emph{photosynthesis})
			\item High dead moisture fuel content (\emph{humidity})
			\item Poor convection/smoke dispersal (\emph{humidity})
		\end{itemize}
		\item[]
		\item Social 
		\item[] \emph{mostly, \alert{too dry}}
		\begin{itemize}
			\item Local burn restrictions
			\item Control issues
		\end{itemize}
	\end{itemize}

@@

\begin{frame}{What a difference a few days make! } 
	\begin{columns}
		\begin{column}{0.5\textwidth}
				\begin{center}
				\begin{figure}
					%\caption{5 May 2018}
					\includegraphics[width=0.9\linewidth]{figs/GoodBurn} 
				\end{figure}
			\end{center}
		\end{column}
		\begin{column}{0.5\textwidth}  
			\begin{center}
				\begin{figure}
					%\caption{16 May 2018}
					\includegraphics[width=1\linewidth]{figs/StreeterFire_17}  
				\end{figure}
			\end{center}
		\end{column}
	\end{columns}
\centering
\adjustbox{max height=\dimexpr\textheight-5.5cm\relax,
	max width=1\textwidth}{
	\begin{tabular}{ccc}
		\toprule
		5 May 2018 & \textbf{Date} & 16 May 2018 \\
		79 & \textbf{Air temp (F)} & 82 \\
		5.9 & \textbf{Wind (m s\textsuperscript{-1})} & 2.4 \\
		22 & \textbf{min RH (\%)} & 24 \\
		37 & \textbf{Dew Point (F)} & 46 \\\bottomrule
	\end{tabular}
}
\end{frame}

@@

\begin{frame}{The terrible, humid, green, no-good day}

	\begin{columns}
		\begin{column}{0.6\textwidth}
	\alert{13 August 2018}
	\begin{itemize} 
		\item Forecast called for $<~40\%$ RH by noon, but...
		\begin{itemize}
			\item 54\% at 1200
			\item 50\% at 1500
		\end{itemize}
	\item[]
		\item Modeled historical data (gridMET) says RH reached 28\%
	\item[]
		\item Local factors constrain ignition and spread
				\begin{itemize}
			\item Transpiration by green plants
			\item Humid air resists heating, lift
		\end{itemize}

	\end{itemize}

		\end{column}
		\begin{column}{0.4\textwidth}  
			\begin{center}
				\begin{figure}
					\includegraphics[width=0.9\linewidth]{figs/TorreGreenFuel}

@@

\begin{frame}{Fire regime management}
	Difficult to satisfy multiple socially-constructed parameters (e.g. \emph{force patterns})
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/venn} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{Fire regime management}
	\vspace{-3em}
	Better to tweak controllable components towards desired objectives (e.g. \emph{support processes})
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/levers} 
		\end{figure}
	\end{center}
\end{frame}

\begin{frame}{Fire regime management}
	Better to tweak controllable components towards desired objectives (e.g. \emph{support processes})
	\begin{center}
		\begin{figure}
			\includegraphics[width=1\linewidth]{figs/levers} 
		\end{figure}
	\end{center}
\alert{Great Plains are in a \textbf{fire deficit} and management is best focused on \textbf{burning new acres}}
\end{frame}

\begin{frame}{Mopping up}

\begin{columns}
\begin{column}{0.55\textwidth}
\begin{itemize}
	\item Fight the deficit: Burn what you can, when you can, but try to add new acres
	\item[]
		\item Focus on fine-scale levers to accomplish objectives regardless of season
	\item[]
	\item Consider emphasis on burn completeness: \emph{defend desired refugia}

\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}  
\begin{center}
\begin{figure}
\includegraphics[width=1\linewidth]{figs/mop_up} 
 \end{figure}
\end{center}
\end{column}
\end{columns}
\end{frame}
